---
title: "FinalProject.DataAnalysis"
author: "Sun Park"
date: "12/2/2020"
output:
  html_document:
    df_print: paged
---

```{r setup}
library(psych)
library(plyr)
naas2016 <- read.csv("/Users/hyunsunpark/Desktop/QMSS_FL20/Data Analysis/NAAS-data.csv")
naas2016[naas2016 == 88 | naas2016 == 99] <- NA
naas2016apia <- subset(naas2016, RACE < 3)
table(naas2016apia$RACE)
```

```{r recodes}
naas2016apia$american <- ifelse(naas2016apia$Q4_101 == 5, 1, 0) ## if identifies as solely American = 1, other identifications = 0
table(naas2016apia$american)
naas2016apia$asianid <- naas2016apia$Q4_2A ## How important is being ASIAN to your identity?
table(naas2016apia$asianid)
naas2016apia$sq.asianid <- naas2016apia$asianid * naas2016apia$asianid
naas2016apia$promote <- ifelse(naas2016apia$Q5_2_A == 1, 1, 0) ## if unfairly denied promotion = 1, if not = 0
table(naas2016apia$promote)
naas2016apia$name <- ifelse(naas2016apia$Q5_1_I == 1, 1, 0) ## if people mispronounce name = 1, if not = 0
table(naas2016apia$name)
naas2016apia$mathsci <- ifelse(naas2016apia$Q5_1_J == 1, 1, 0) ## if people assume respondent good at math and science = 1, if not = 0 
table(naas2016apia$mathsci)
naas2016apia$amerid <- naas2016apia$Q4_2D ## How important is being AMERICAN to your identity?
table(naas2016apia$amerid)
naas2016apia$female <- ifelse(naas2016apia$S7 == 2, 1, 0)
table(naas2016apia$female)

naas2016apia$affect <- 4 - naas2016apia$Q4_3A ## what happens to Asian/Asian-Americans affect personally 
table(naas2016apia$affect)
naas2016apia$commcul <- ifelse(naas2016apia$Q4_5B == 1, 1, 0) ## Does America share a common culture? 
table(naas2016apia$commcul)


## variables that are currently not being used
# affect ; commcul ; 
```

## Descriptive Statistics 
```{r descriptive stats}
describe(naas2016apia$american)
describe(naas2016apia$asianid)
describe(naas2016apia$promote)
describe(naas2016apia$amerid)
describe(naas2016apia$mathsci)
describe(naas2016apia$name)
describe(naas2016apia$female)
```

## Initial regression 
```{r model promote}
american.lm <- lm(promote ~ american, naas2016apia)
summary(american.lm)
```

## Initial regression with only American Identification 
```{r model amerid}
amerid.lm <- lm(promote ~ american + amerid, naas2016apia)
summary(amerid.lm)
```

## Initial regression with only Asian Identification 
```{r model asianid}
asianid.lm <- lm(promote ~ american + asianid, naas2016apia)
summary(asianid.lm)
```

## Second regression with both American and Asian Identification  
```{r model amerid asianid}
aaid.lm <- lm(promote ~ american + amerid + asianid, naas2016apia)
summary(aaid.lm)
```

## Third regression with both American and Asian Identification as interaction
```{r model amerid:asianid}
aaidint.lm <- lm(promote ~ american + amerid*asianid + amerid + asianid  + female, naas2016apia)
summary(aaidint.lm)
```

```{r plot}
library(ggplot2)
plot_aaid <- ggplot(naas2016apia, aes(x = asianid, y = promote, group = amerid, color = amerid))
plot_aaid + stat_smooth(method = "lm", se = F)
```

## Fourth regression with interaction of identifications and Asian stereotypes
```{r model stereo}
stereo.lm <- lm(promote ~ american + asianid:amerid + name + mathsci + asianid + amerid + female, naas2016apia)
summary(stereo.lm)
```

```{r anova test}
anova(aaidint.lm, stereo.lm$model)
```


## Fifth regression as logistic regression
```{r logit}
asianid.lgt <- glm(promote ~ american + asianid, naas2016apia, family = binomial)
summary(asianid.lgt)
```

## Sixth regression as logistic regression with additional predictors
```{r logit stereo}
stereo.lgt <- glm(promote ~ american + asianid + amerid + female + asianid:amerid + name + mathsci, naas2016apia, family = binomial)
summary(stereo.lgt)
```

```{r anova}
anova(stereo.lgt, test = "Chisq")
library(pscl)
pR2(stereo.lgt)
```
